---
title: "Uneven Investments in Toronto's Wards"
subtitle: "An Analysis of Capital Funding Allocation in Toronto's Wards from 2022 to 2031"
author: Maria Mangru 
thanks: "Code and data are available at: https://github.com/MariaMangru/Capital-Budget-Allocations-and-Median-Household-Income."
date: today
date-format: long
abstract: "This study examines the allocation of capital funding across Toronto's 25 wards from 2022 to 2031, focusing on its relationship with key socioeconomic factors. The results found that wards with higher median household incomes and greater educational attainment receive more captial funding per capita. Conversely, wards with higher proportions of low-income households and visible minority populations tend to receive less funding per capita."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Toronto is a city renowned for its cultural diversity, however, it is also a city marked by stark socioeconomic disparities. Each year, Toronto allocates funds through its capital budget to support essential services such as public transit, libraries, parks and water facilities. \[REF: https://www.toronto.ca/wp-content/uploads/2017/11/97f7-A170XXXX_Budget_Basics_Understanding-final-web.pdf\]. This funding shapes the city's infrastructure and directly impacts residents' quality of life and access to opportunities.

Despite Toronto's commitment to equity, evidence suggests that wealthier neighborhoods receive a disproportionate share of public investment. Previous reports, such as Walks et al. (2016) \[REFERENCE HERE\], highlight how income segregation leads to affluent areas accumulating more resources, while low-income neighborhoods face underfunded services. This pattern enhances socioeconomic disparities and hinders the city's overall progress toward inclusivity.

This research investigates how capital funding from 2022 to 2031 is allocated across Toronto's wards and examines its relationship with median household income, educational attainment, and visible minority populations. By identifying patterns in funding distribution, we aim to understand whether current investment strategies align with the needs of all communities.

The remainder of the research is organized as follows: Section \ref{sec-data} describes the data sources and methodology used in the analysis. Section \ref{sec-results} presents the results, highlighting key findings through visualizations and statistical summaries. Section \ref{sec-discussion} discusses the implications of these findings, addresses limitations, and offers recommendations for further research.

# Data {#sec-data}

## Data Sources

This study utilizes two primary datasets provided by the City of Toronto's Open Data Portal [REFERENCE].

1.  Capital Budget & Plan By Ward (2022-2031):

This dataset provides detailed funding allocations for each of Toronto's 25 wards over a ten-year period It covers various programs such as Children's Services, Parks, Forestry & Recreation, and Transportation Services. The budget amounts are in thousands of dollars where a value of 10 represents \$10,000. Additionally, the dataset includes both immediate five-year budgets (Year 1 to Year 5) and subsequent five-year planned budgets (Year 6 to Year 10), with a cumulative total for the ten-year span.

2.  Ward Profiles (2021):

Based on the 25-Ward model effective from December 1, 2018, this dataset provides comprehensive socioeconomic information derived from the 2021 Census. It includes metrics such as total population, median household income, prevalence of low-income households, educational attainment, and visible minority populations.

These datasets were merged in order to link capital funding allocations with socioeconomic indicators at the ward level. The analysis for this paper was conducted using @citeR, with several packages from the @tidyverse suite being used for data cleaning, manipulation and visualization. Visualizations were created using @ggplot2, and tables were rendered using kableExtra [REFERENCE].


### Data Collection and Measurement 

The Capital Budget & Plan By Ward dataset is manually updated and published annually by the City of Toronto, capturing both approved and planned funding across various programs and projects within each ward. Key attributes include program or agency names, project and sub-project names, annual budget allocations from Year 1 to Year 10, total ten-year budgets, ward numbers and names, and project categories such as infrastructure or community services.

The Ward Profiles dataset includes a wide range of demographic, social, and economic indicators such as population, household information, migration and mobility patterns, indigenous and ethnocultural composition, language and education levels, as well as labor force and income statistics. It should be acknowledged that census data may be subject to sampling and non-response biases, though adjustments are made to mitigate these issues. 

Both datasets present certain limitations. The Capital Budget & Plan By Ward dataset’s manual updates introduce potential for human error and delays in reflecting the most current budget allocations. Similarly, the Ward Profiles dataset excludes data from the 2011 National Household Survey, limiting longitudinal analyses beyond a decade. Adjustments made to align older Census data with current ward boundaries may also introduce inaccuracies or inconsistencies. 


```{r setting up workspace, echo=FALSE, message=FALSE, warning=FALSE}
#### Workspace setup ####
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
library(kableExtra)
library(scales)
library(ggplot2)
library(stringr)
library(janitor)


### Data Handling ###

# Read in the cleaned budget data
budget_data <- read_csv("../data/analysis_data/clean_budget_data.csv")

# Read in the cleaned demographic data
demographic_data <- read_csv("../data/analysis_data/clean_ward_demo_data.csv")

# Identify year columns in budget data
year_cols <- as.character(2022:2031)

# Aggregate total funding per ward number
total_funding_ward <- budget_data |>
  group_by(`Ward Number`)|>
  summarise(
    total_funding = sum(across(all_of(year_cols)), na.rm = TRUE)
  ) |>
  arrange(desc(total_funding))

# Adjust total funding (amounts are in thousands of dollars)
total_funding_ward <- total_funding_ward |>
  mutate(total_funding = total_funding * 1000)

# Pivot demographic data from wide to long format
demographic_long <- demographic_data |>
  pivot_longer(
    cols = -category,
    names_to = "Ward Number",
    values_to = "Value"
  )

# Pivot demographic data back to wide format to get one row per ward
demographic_wide <- demographic_long |>
  pivot_wider(
    names_from = category,
    values_from = Value
  )

# Convert 'Ward Number' to numeric in both datasets to allow merging
total_funding_ward$`Ward Number` <- as.numeric(as.character(total_funding_ward$`Ward Number`))
demographic_wide$`Ward Number` <- as.numeric(as.character(demographic_wide$`Ward Number`))

# Merge budget data with demographic data
merged_data <- total_funding_ward |>
  left_join(demographic_wide, by = "Ward Number")

# Convert 'Total Population' to numeric
merged_data$`Total Population` <- as.numeric(gsub(",", "", merged_data$`Total Population`))


# Create per capita funding variable
merged_data <- merged_data |>
  mutate(per_capita_funding = total_funding / `Total Population`)


# Convert columns from list to double 
merged_data$`Bachelor’s degree or higher` <- as.numeric(merged_data$`Bachelor’s degree or higher`)
merged_data$`Total visible minority population` <- as.numeric(merged_data$`Total visible minority population`)


#Create per capita bachelor's degree variable
merged_data <- merged_data |>
  mutate(`Bachelor’s degree or higher per capita` = `Bachelor’s degree or higher` / `Total Population`)


#Create per capita visible minority population variable
merged_data <- merged_data |>
  mutate(`Total visible minority population per capita` = `Total visible minority population` / `Total Population`)

# Create a data frame with Ward Names 
ward_names_df <- data.frame(
  `Ward Number` = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 
                    11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 
                    21, 22, 23, 24, 25),
  `Ward Name` = c("Etobicoke North", "Etobicoke Centre", "Etobicoke—Lakeshore", "Parkdale—High Park", "York South—Weston", "York Centre",
                "Humber River—Black Creek", "Eglinton—Lawrence", "Davenport", "Spadina—Fort York", "University—Rosedale", "Toronto—St. Paul's", "Toronto Centre", "Toronto—Danforth", "Don Valley West", "Don Valley East", "Don Valley North", "Willowdale", "Beaches—East York", "Scarborough Southwest", "Scarborough Centre", "Scarborough—Agincourt", "Scarborough North", "Scarborough Guildwood", "Scarborough—Rouge Park"),
  check.names = FALSE
)

# Join ward names into marged data
merged_data <- merged_data |>
  left_join(ward_names_df, by = "Ward Number")

```

## Data Preparation

This research focused on ward-specific funding by excluding "City Wide" allocations that are not tied to particular wards. The per capita funding was calculated by dividing each ward's total funding by its population. Similarly, the variables `Educational Attainment` and `Visible Minority Population` were were expressed as percentages by dividing each variable by the ward's population. This was done to standardized the variables allowing the comparison across wards, ensuring that variations in population sizes do not skew the analysis.

The key variables of interest for this research are:

-   Per Capita Funding: Total capital funding allocated to a ward divided by its population.

-   Median Household Income: The median income of households within a ward.

-   Low-Income Prevalence: The percentage of households classified as low-income in a ward.

-   Educational Attainment: The percentage of residents with a bachelor's degree or higher.

-   Visible Minority Population: The percentage of residents who identify as visible minorities.


Although multiple ten-year budget datasets are available, including the 2023-2032 and 2024-2033 Capital Budget & Plan, this study opted to utilize the 2022-2031 Capital Budget & Plan. This is because, the 2022-2031 budget period aligns more closely with the 2021 Ward Profiles, ensuring that the socioeconomic variables and funding allocations are analyzed within a similar timeframe. 

```{r summary-stats-table, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-demographics-stats
#| tbl-cap: Table Showing Capital Funding and Socioeconomic Indicators by Ward

# Create a combined Ward column
merged_data <- merged_data |>
  mutate(`Ward` = paste(`Ward Name`, "-", `Ward Number`))


# Select relevant columns for the table
merged_table_select <- merged_data |>
  select(
    `Ward`, 
    per_capita_funding,
    `Median total income of households in 2020 ($)`,
    `Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`,
    `Bachelor’s degree or higher per capita`,
    `Total visible minority population per capita`
  )

# Arrange by the numeric version of per capita funding before formatting
merged_table_select <- merged_table_select |>
  arrange(desc(per_capita_funding))

# Format numeric columns with conditional formatting for LaTeX
merged_table_2 <- merged_table_select |>
  mutate(
    # Format Per Capita Funding as dollars
    per_capita_funding = dollar(per_capita_funding),
    `Median total income of households in 2020 ($)` = paste0(dollar(as.numeric(`Median total income of households in 2020 ($)`))),
    `Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)` = paste0(round(as.numeric(`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`) * 100, 2), "%"),
    `Bachelor’s degree or higher per capita` = paste0(round(as.numeric(`Bachelor’s degree or higher per capita`) * 100, 2), "%"),
    `Total visible minority population per capita` = paste0(round(as.numeric(`Total visible minority population per capita`) * 100, 2), "%")
  )

# Create the table with kable and apply enhanced styling
kable(
  merged_table_2, 
  escape = TRUE,                   
  col.names = c(
    "Ward Name - Number", 
    "Per Capita Funding", 
    "Household Median Income", 
    "Low Income Prevalence", 
    "Bachelor’s Degree or Higher", 
    "Visible Minority Population"
  )
) |>
  kable_styling(
    bootstrap_options = c("condensed"),  
    full_width = FALSE,  
    position = "left",  
    font_size = 7 
  ) |>
  footnote(general =  "Rows in red indicate wards with a Household Median Income at or below the average of $84,000.") |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "3cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5:6, width = "3cm") |>
  row_spec(1, color = "red") |>
  row_spec(6, color = "red") |>
  row_spec(9:10, color = "red") |>
  row_spec(13:14, color = "red") |>
  row_spec(17:19, color = "red") |>
  row_spec(21:24, color = "red")
```


@tbl-demographics-stats shows the ward specific funding and key demographic indicators. Ward 13 stands out with the highest per capita funding and a high percentage of low income prevalence, relative to other wards.

\newpage

```{r Figure 1, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-percapitafunding
#| fig-cap: Bar graph showing Per Capita Capital Funding by Ward

# Wards by Per Capita Funding
ggplot(merged_data, aes(x = reorder(`Ward`, per_capita_funding), y = per_capita_funding)) +
  geom_bar(stat = "identity", fill = "#0072B2") +
  labs(
    title = "Wards by Per Capita Funding",
    x = "Ward Name - Number",
    y = "Per Capita Funding ($CAD)"
  ) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 14),
    axis.text.y = element_text(size = 10)
  ) +
  coord_flip()
```

@fig-percapitafunding gives a visual representation of the per capita funding allocated to each ward. This per capita funding varied widely, ranging from \$173.22 in Ward 15 to \$7,674.47 in Ward 13. This significant variation indicates an uneven distribution of resources which may not align with the needs of all communities.

# Results {#sec-results}

## Per Capita Funding and Household Median Income

The average income across wards is \$84,000, meaning half of the wards have median incomes at or below this amount. Higher income wards will be categorized as one with a median household income above \$84,000. A positive relationship was observed between per capita funding and median household income as wards above the citywide average \$84,000 generally received more funding per capita. Specifically, among the top five wards in per capita funding, four have median incomes above \$84,000. However, among the bottom five wards, four have median incomes below \$84,000.

More specifically, Ward 2, with a high median income of \$100,000, received the second-highest per capita funding at \$4,166.43. In contrast, Ward 22, which has a lower median income of \$77,000, received the second-lowest per capita funding at just \$192.44. This pattern suggests that wealthier wards benefit more from capital investments.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-householdincome
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Median Household Income

# Scatter plot of Per Capita Funding vs Median Household Income with ward 13 removed 

filtered_data <- merged_data |>
  filter(`Ward Number` != 13)

ggplot(filtered_data, aes(x = as.numeric(`Median total income of households in 2020 ($)`), y = per_capita_funding)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Per Capita Funding vs Median Household Income",
    x = "Median Household Income ($)",
    y = "Per Capita Funding ($)"
  ) +
  scale_x_continuous(labels = dollar_format()) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal()
```

@fig-householdincome shows the relationship between the variables, with Ward 13 excluded from the analysis due to its disproportionate impact, which distorts the overall data and does not provide an accurate depiction of the relationship. The upward trend line indicates a positive relationship between per capita funding and median household income. In other words, as median household income rises, per capita funding tends to increase as well.

## Per Capita Funding and Low-Income Prevalence

It was found that an inverse relationship exists between per capita funding and the prevalence of low income households. We will classify a ward as having a high percentage of low-income households if the prevalence is 13% or higher, which represents the average low-income prevalence. Among the top five funded wards, only two can be classified as higher low-income prevalence. In contrast, of the bottom five funded wards, four can be classified as having higher low-income prevalence. This suggests that wards with more low-income households receive less capital budget funding.

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-lowincome
#| fig-cap: Per capita funding and low income prevalence for top and bottom 5 wards excluding ward 13 

top5_wards <- merged_data |>
  arrange(desc(per_capita_funding)) |>
  slice(2:6)

bottom5_wards <- merged_data |>
  arrange(per_capita_funding)|>
  slice(1:5)

# Combine these wards 
selected_wards <- bind_rows(top5_wards, bottom5_wards) |>
  mutate(Category = ifelse(`Ward` %in% top5_wards$`Ward`, "Top 5", "Bottom 5"))

selected_wards$`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)` <- as.numeric(as.character(selected_wards$`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`))


# Calculate scaling factor
max_funding <- max(selected_wards$per_capita_funding, na.rm = TRUE)
max_low_income <- max(selected_wards$`Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)`, na.rm = TRUE)
scale_factor <- max_funding / max_low_income

# Add a scaled Low Income Prevalence for plotting
selected_wards <- selected_wards |>
  mutate(Scaled_Low_Income_Prevalence = `Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%)` * scale_factor)


# Create the Combined Bar and Line Chart
ggplot(selected_wards, aes(x = reorder(`Ward`, per_capita_funding))) +
  geom_bar(aes(y = per_capita_funding, fill = Category), stat = "identity") +
  geom_line(aes(y = Scaled_Low_Income_Prevalence, group = 1, color = "Low Income Prevalence"), size = 1.2) +
  geom_point(aes(y = Scaled_Low_Income_Prevalence, color = "Low Income Prevalence"), size = 3) +
 
   scale_y_continuous(
    name = "Per Capita Funding ($CAD)",
    labels = dollar_format(),
    sec.axis = sec_axis(~./scale_factor, name = "Low Income Prevalence (Decimal)")
  ) +
  labs(
    title = "Top 5 and Bottom 5 Wards by Per Capita Funding with Low Income Prevalence",
    x = "Ward Name - Number",
    fill = "Category",
    color = ""
  ) +
  

  scale_fill_manual(values = c("Top 5" = "#0072B2", "Bottom 5" = "red")) +
  scale_color_manual(values = c("Low Income Prevalence" = "black")) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.y.left = element_text(color = "black", size = 10),
    axis.title.y.right = element_text(color = "black", size = 10),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8)
  )
```

@fig-lowincome compares the top and bottom 5 wards by per capita funding, with the prevalence of low-income households overlaid. The red bars represent the wards with the lowest per capita funding, while the blue bars represent those with the highest. Ward 13 was excluded from this graph due to its skewed impact on the data. The black line shows the prevalence of low-income households, which is generally higher in wards with lower per capita funding. However, even among the top-funded wards, the relationship between per capita funding and low-income prevalence is not perfectly linear, suggesting that while funding tends to be higher in wealthier wards, some lower-income wards still receive significant funding.

## Per Capita Funding and Educational Attainment

We will also classify wards where at least 33% of residents hold a bachelor's degree or higher as having high educational attainment. There is a clear positive relationship between educational attainment and per capita funding as wards with higher education levels tend to receive more funding. More specifically, Ward 10, with 60.97% of residents holding a bachelor’s degree, received \$4,127.10 per capita, while Ward 11, with 56.07%, received \$1,637.81 per capita. In contrast, Ward 7, where only 14.67% of residents hold a degree, received just \$852.05 per capita.

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-education
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Education Levels

# Scatter plot of Per Capita Funding vs Bachelor's Degree of Higher with ward 13 removed 

filtered_data <- merged_data |>
  filter(`Ward Number` != 13)

ggplot(filtered_data, aes(x = as.numeric(`Bachelor’s degree or higher per capita`), y = per_capita_funding)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Per Capita Funding vs Bachelor's degree or higher",
    x = "Attained Bachelor's degree or higher (%)",
    y = "Per Capita Funding ($)"
  ) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal()
```

The trend line in @fig-education illustrates that, generally, wards with a higher percentage of degree holders are more likely to see higher per capita funding allocations. Ward 13 was excluded from this analysis due to its skewing effect on the data.

## Per Capita Funding and Visible Minority Population

The relationship between per capita funding and the proportion of visible minority populations appears complex, with signs of potential disparities. Wards with higher visible minority populations often receive less funding per capita. We will consider a ward to have a high percentage of visible minorities if 57% or more of its population falls into this category. 57% represents the average percent of visible minorities across the wards. Among the top ten funded wards, only four have higher proportions of visible minorities, while among the bottom ten, six wards have higher visible minority populations.

Examples highlighting this trend include Ward 22, where 82.31% of the population are visible minorities, and which received only \$192.44 per capita, and Ward 23, with 92.18% visible minorities, which received \$1,829.15 per capita. In contrast, Ward 2, with a lower visible minority population of 31.75%, received significantly more funding at \$4,166.43 per capita. While exceptions exist, the general trend suggests that wards with larger visible minority populations may receive less funding.

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-minority
#| fig-cap: Scatter plot illustrating the relationship between Per Capita Funding and Minority Population

# Scatter plot of Per Capita Funding vs Minority Population with ward 13 removed 

filtered_data <- merged_data |>
  filter(`Ward Number` != 13)

ggplot(filtered_data, aes(x = as.numeric(`Total visible minority population per capita`), y = per_capita_funding)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Per Capita Funding vs Visible Minority Population",
    x = "Visible Minority Population (%)",
    y = "Per Capita Funding ($)"
  ) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal()
```

@fig-minority illustrates a downward trend between per capita funding and the percentage of visible minority populations across wards, with Ward 13 excluded. As the visible minority population increases, per capita funding tends to decrease. This trend suggests potential disparities in how funding is distributed among wards with higher visible minority populations.

# Discussion {#sec-discussion}

This study's examination of capital funding allocation across Toronto's wards reveals patterns that mirror broader socioeconomic dynamics discussed in urban studies literature. The disparities observed align with the findings of Walks et al. (2016) \[REFERENCE\], who highlight how income segregation in Canadian cities leads to affluent areas accumulating more resources while low-income neighborhoods face underfunded services. These patterns contribute to socio-spatial income inequality, where wealth and resources are unevenly distributed, reinforcing existing socioeconomic divides.

The unequal distribution of capital funding has significant implications for both present and future generations. Underfunded communities often lack access to essential services such as quality education, healthcare, and public transportation. The National Academies of Sciences, Engineering, and Medicine (2023) \[REFERENCE\] emphasize that children growing up in low-income households are more likely to experience intergenerational poverty, which can persist without targeted interventions. Limited access to resources in underfunded wards can hinder social mobility and perpetuate cycles of poverty.

Furthermore, the disparities in funding contribute to a widening wealth gap within the city. Wealthier wards receiving more investment can improve infrastructure, increase property values, and attract further economic development, as discussed by Walks et al. (2016) \[REFERENCE\]. In contrast, underfunded wards may experience deteriorating infrastructure and services, making them less attractive for investment and leading to economic stagnation. This dynamic exacerbates socioeconomic disparities and can strain social services, as residents in underfunded areas may rely more heavily on social assistance programs due to a lack of access to resources.

These findings highlight the broader effects of unequal resource distribution on urban communities. The reinforcement of existing inequalities through funding allocation can lead to increased social stratification, reduced social cohesion, and limited opportunities for residents in disadvantaged areas. Addressing these disparities requires a comprehensive understanding of how funding decisions impact different communities and the implementation of strategies that promote equitable access to resources.

## Weaknesses and next steps

### Weaknesses

While this study provides interesting results, there are a some limitations. Firstly, the analysis is limited by the focus on median household income, educational attainment, and visible minority populations, potentially missing other important socioeconomic factors like housing costs, employment rates, and age demographics. Additionally, the reliance on 2021 census data and projected capital budgets through 2031 may not account for recent changes. Lastly, the methodology highlights associations without examining the underlying causes of funding disparities, such as political priorities or historical patterns.

### Next Steps

To address these limitations, future research should incorporate a broader range of socioeconomic variables, such as housing costs and employment rates, to gain a more comprehensive understanding of funding distribution. Comparative analyses with other cities could also help identify best practices for equitable funding and inform more effective policy decisions. Lastly, conducting a longitudinal study of multiple capital budgets over time would be beneficial in evaluating long-term trends in capital investments. 

\newpage

# References
